---
title: "Rsassy, sassyRS, and Python binding benchmark"
output:
  html_document:
    toc: true
    toc_depth: 2
  github_document:
    toc: true
    toc_depth: 2
params:
  batches: 7
  iterations: 20
  text_length: 1000000
  warmup: 2
  output: "r/benchmarks/results/r-bindings.csv"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")

library(reticulate)

param_int <- function(env, default) {
  value <- Sys.getenv(env, unset = "")
  if (nzchar(value)) as.integer(value) else as.integer(default)
}

find_repo_root <- function(path = getwd()) {
  path <- normalizePath(path, mustWork = TRUE)
  while (!file.exists(file.path(path, "Cargo.toml"))) {
    parent <- dirname(path)
    if (identical(parent, path)) stop("Could not find repository root")
    path <- parent
  }
  path
}

repo_root <- find_repo_root()
batches <- param_int("SASSY_R_BENCH_BATCHES", params$batches)
iterations <- param_int("SASSY_R_BENCH_ITERS", params$iterations)
text_length <- param_int("SASSY_R_BENCH_TEXT_LENGTH", params$text_length)
warmup <- param_int("SASSY_R_BENCH_WARMUP", params$warmup)
output_csv <- Sys.getenv("SASSY_R_BENCH_OUT", params$output)
if (!grepl("^/|^[A-Za-z]:[/\\\\]", output_csv)) output_csv <- file.path(repo_root, output_csv)
```

This report compares the R-facing overhead of three bindings to `sassy`:

- **Rsassy**: R bindings through R's native C API.
- **sassyRS**: an `extendr` prototype.
- **Python sassy via reticulate**: the published `sassy-rs` Python binding called from R with `reticulate`.

The benchmark uses a deterministic 1 Mbp DNA text by default, not the tiny toy string used in examples. Exact and two-mismatch copies of the query are inserted every 100 kbp so each binding has real matches to return.

## Machine and build details

The native benchmark build is expected to be installed before rendering with:

```bash
RUSTFLAGS="-C target-cpu=native" SASSY_RUST_FEATURES=native-simd R CMD INSTALL r/Rsassy
RUSTFLAGS="-C target-cpu=native" SASSY_RUST_FEATURES=native-simd R CMD INSTALL r/sassyRS
uv venv --python 3.11 .venv-benchmark
uv pip install --python .venv-benchmark sassy-rs
export RETICULATE_PYTHON="$PWD/.venv-benchmark/bin/python"
```

```{r machine, echo=FALSE}
cmd <- function(...) paste(system2(..., stdout = TRUE, stderr = TRUE), collapse = "\n")
lscpu <- system2("lscpu", stdout = TRUE, stderr = TRUE)
cpu_model <- sub("^Model name:[[:space:]]*", "", grep("^Model name:", lscpu, value = TRUE)[1])
cpu_flags <- sub("^Flags:[[:space:]]*", "", grep("^Flags:", lscpu, value = TRUE)[1])
simd_flags <- paste(grep("^(sse|avx|fma|bmi)", strsplit(cpu_flags, "[[:space:]]+")[[1]], value = TRUE), collapse = " ")
python_bin <- Sys.getenv("RETICULATE_PYTHON", "python3")

machine <- data.frame(
  field = c(
    "system", "release", "machine", "cpu_model", "R", "platform",
    "RUSTFLAGS", "SASSY_RUST_FEATURES", "RETICULATE_PYTHON", "cargo", "rustc", "python", "simd_flags"
  ),
  value = c(
    unname(Sys.info()[["sysname"]]),
    unname(Sys.info()[["release"]]),
    unname(Sys.info()[["machine"]]),
    cpu_model,
    R.version.string,
    R.version$platform,
    Sys.getenv("RUSTFLAGS", unset = ""),
    Sys.getenv("SASSY_RUST_FEATURES", unset = ""),
    Sys.getenv("RETICULATE_PYTHON", unset = ""),
    cmd("cargo", "--version"),
    cmd("rustc", "--version"),
    cmd(python_bin, "--version"),
    simd_flags
  )
)
knitr::kable(machine)
```

## Input data

```{r input, echo=FALSE}
mutate_base <- function(x) switch(x, A = "C", C = "G", G = "T", T = "A")
mutate_pattern <- function(pattern, positions = c(11L, 29L)) {
  chars <- strsplit(pattern, "", fixed = TRUE)[[1]]
  chars[positions] <- vapply(chars[positions], mutate_base, character(1))
  paste(chars, collapse = "")
}

make_text <- function(pattern, n, spacing = 100000L) {
  unit <- paste0(
    "TGCATGATCGATACGTTAGCCTAGGCTAACCGATCGGCTA",
    "CGTATCGGATCCGATGCAATCGTAGCTAGCATCGATGCAT"
  )
  text <- substr(paste(rep(unit, ceiling(n / nchar(unit))), collapse = ""), 1L, n)
  starts <- seq.int(1000L, n - nchar(pattern) + 1L, by = spacing)
  mutated <- mutate_pattern(pattern)
  for (i in seq_along(starts)) {
    substr(text, starts[[i]], starts[[i]] + nchar(pattern) - 1L) <- if (i %% 2L) pattern else mutated
  }
  list(text = text, inserted = length(starts), spacing = spacing)
}

pattern <- "GTACAGAAACGAGCGGATGGAAAGAGTAGTGAGCGCCTCGCG"
k <- 2L
text_spec <- make_text(pattern, text_length)
text <- text_spec$text
raw_pattern <- charToRaw(pattern)
raw_text <- charToRaw(text)

config <- data.frame(
  batches = batches,
  iterations_per_batch = iterations,
  warmup = warmup,
  text_length = nchar(text, type = "bytes"),
  pattern_length = nchar(pattern, type = "bytes"),
  k = k,
  inserted_matches = text_spec$inserted,
  insertion_spacing = text_spec$spacing
)
knitr::kable(config)
```

## Benchmarked calls

```{r benchmark, echo=FALSE}
py_sassy <- import("sassy", convert = FALSE)
py_builtins <- import_builtins(convert = FALSE)
py_pattern <- py_builtins$bytes(raw_pattern)
py_text <- py_builtins$bytes(raw_text)
py_searcher <- py_sassy$Searcher("dna", rc = FALSE)

rsassy_searcher_chr <- Rsassy::sassy_searcher("dna", rc = FALSE)
rsassy_searcher_raw <- Rsassy::sassy_searcher("dna", rc = FALSE)

cases <- list(
  list(
    binding = "Rsassy reusable character",
    run = function() Rsassy::sassy_searcher_search(rsassy_searcher_chr, pattern, text, k),
    count = nrow
  ),
  list(
    binding = "Rsassy reusable raw",
    run = function() Rsassy::sassy_searcher_search(rsassy_searcher_raw, raw_pattern, raw_text, k),
    count = nrow
  ),
  list(
    binding = "Rsassy one-shot raw",
    run = function() Rsassy::sassy_search(raw_pattern, raw_text, k, alphabet = "dna", rc = FALSE),
    count = nrow
  ),
  list(
    binding = "sassyRS one-shot raw",
    run = function() sassyRS::sassy_search(raw_pattern, raw_text, k, alphabet = "dna", rc = FALSE),
    count = nrow
  ),
  list(
    binding = "Python sassy via reticulate",
    run = function() py_searcher$search(py_pattern, py_text, as.integer(k)),
    count = reticulate::py_len
  )
)

bench_case <- function(case) {
  for (i in seq_len(warmup)) case$run()
  gc()

  seconds <- numeric(batches)
  matches <- NA_integer_
  for (batch in seq_len(batches)) {
    last <- NULL
    elapsed <- system.time({
      for (i in seq_len(iterations)) last <- case$run()
    })[["elapsed"]]
    matches <- case$count(last)
    seconds[[batch]] <- elapsed / iterations
  }

  data.frame(
    binding = case$binding,
    batches = batches,
    iterations_per_batch = iterations,
    text_length = nchar(text, type = "bytes"),
    pattern_length = nchar(pattern, type = "bytes"),
    k = k,
    matches = matches,
    min_seconds = min(seconds),
    median_seconds = unname(stats::median(seconds)),
    mean_seconds = mean(seconds),
    p95_seconds = unname(stats::quantile(seconds, 0.95, names = FALSE)),
    max_seconds = max(seconds),
    calls_per_second = 1 / unname(stats::median(seconds))
  )
}

results <- do.call(rbind, lapply(cases, bench_case))
results <- results[order(results$median_seconds), ]
row.names(results) <- NULL
knitr::kable(results)
```

## Summary

```{r summary, echo=FALSE}
summary_table <- results[, c("binding", "median_seconds", "calls_per_second", "matches")]
knitr::kable(summary_table)

dir.create(dirname(output_csv), recursive = TRUE, showWarnings = FALSE)
utils::write.csv(results, output_csv, row.names = FALSE)
message("Wrote benchmark results to ", output_csv)

summary_path <- Sys.getenv("GITHUB_STEP_SUMMARY", unset = "")
if (nzchar(summary_path)) {
  lines <- c(
    "## Rsassy, sassyRS, and Python sassy benchmark",
    "",
    sprintf("Text length: `%s` bytes; pattern length: `%s`; k: `%s`.", nchar(text, type = "bytes"), nchar(pattern, type = "bytes"), k),
    sprintf("Build flags: `RUSTFLAGS=%s`; `SASSY_RUST_FEATURES=%s`.", Sys.getenv("RUSTFLAGS", unset = ""), Sys.getenv("SASSY_RUST_FEATURES", unset = "")),
    "",
    "| Binding | Median seconds/call | Calls/second | Matches |",
    "|---|---:|---:|---:|",
    sprintf("| %s | %.6f | %.2f | %d |", summary_table$binding, summary_table$median_seconds, summary_table$calls_per_second, summary_table$matches),
    "",
    sprintf("CSV artifact: `%s`", output_csv)
  )
  writeLines(lines, summary_path, useBytes = TRUE)
}
```
