---
title: "Rsassy vs sassyRS benchmark"
output:
  html_document:
    toc: true
    toc_depth: 2
params:
  install: false
  batches: 7
  iterations: 50
  text_multiplier: 1000
  warmup: 3
  output: "r/benchmarks/results/r-bindings.csv"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This report benchmarks the two R bindings in this repository:

- **Rsassy**: a C/R API boundary intended for explicit raw-vector and future ALTREP access.
- **sassyRS**: an `extendr` boundary prototype.

The benchmark uses a fixed DNA pattern and target text, then reports elapsed seconds per call.

```{r helpers}
get_int_env <- function(name, default) {
  value <- Sys.getenv(name, unset = "")
  if (identical(value, "")) {
    return(as.integer(default))
  }
  value <- suppressWarnings(as.integer(value))
  if (is.na(value) || value <= 0L) {
    stop(name, " must be a positive integer", call. = FALSE)
  }
  value
}

is_absolute_path <- function(path) {
  grepl("^/", path) || grepl("^[A-Za-z]:[/\\\\]", path)
}

find_repo_root <- function(start) {
  path <- normalizePath(start, mustWork = TRUE)
  repeat {
    if (file.exists(file.path(path, "Cargo.toml")) && dir.exists(file.path(path, "r"))) {
      return(path)
    }
    parent <- dirname(path)
    if (identical(parent, path)) {
      stop("Could not locate repository root from ", start, call. = FALSE)
    }
    path <- parent
  }
}

input <- knitr::current_input()
input_dir <- if (!is.null(input) && file.exists(input)) dirname(normalizePath(input)) else getwd()
repo_root <- find_repo_root(input_dir)

batches <- get_int_env("SASSY_R_BENCH_BATCHES", params$batches)
iterations <- get_int_env("SASSY_R_BENCH_ITERS", params$iterations)
text_multiplier <- get_int_env("SASSY_R_BENCH_TEXT_MULTIPLIER", params$text_multiplier)
warmup <- get_int_env("SASSY_R_BENCH_WARMUP", params$warmup)
output <- Sys.getenv("SASSY_R_BENCH_OUT", params$output)
if (!is_absolute_path(output)) {
  output <- file.path(repo_root, output)
}

run <- function(cmd, args) {
  message("$ ", cmd, " ", paste(args, collapse = " "))
  status <- system2(cmd, args)
  if (!identical(status, 0L)) {
    stop("Command failed: ", cmd, call. = FALSE)
  }
}
```

## Package setup

```{r package-setup}
if (isTRUE(params$install)) {
  bench_lib <- tempfile("sassy-r-bindings-lib-")
  dir.create(bench_lib, recursive = TRUE)
  .libPaths(c(bench_lib, .libPaths()))
  message("Installing benchmark packages into ", bench_lib)
  run(file.path(R.home("bin"), "R"), c("CMD", "INSTALL", "-l", bench_lib, file.path(repo_root, "r", "Rsassy")))
  run(file.path(R.home("bin"), "R"), c("CMD", "INSTALL", "-l", bench_lib, file.path(repo_root, "r", "sassyRS")))
}

if (!requireNamespace("Rsassy", quietly = TRUE)) {
  stop("Package 'Rsassy' is not installed. Render with params = list(install = TRUE).", call. = FALSE)
}
if (!requireNamespace("sassyRS", quietly = TRUE)) {
  stop("Package 'sassyRS' is not installed. Render with params = list(install = TRUE).", call. = FALSE)
}
```

## Machine and build details

The benchmark workflow sets `RUSTFLAGS=-C target-cpu=native` and
`SASSY_RUST_FEATURES=native-simd` so the Rust crates are compiled for the
runner CPU instead of the portable scalar fallback used for CRAN-style checks.

```{r machine-details}
safe_system <- function(cmd, args = character()) {
  if (!nzchar(Sys.which(cmd))) {
    return(character())
  }
  out <- tryCatch(system2(cmd, args, stdout = TRUE, stderr = TRUE), error = function(e) character())
  out[!is.na(out)]
}

machine <- data.frame(
  field = c(
    "system",
    "release",
    "machine",
    "R",
    "platform",
    "NOT_CRAN",
    "RUSTFLAGS",
    "SASSY_RUST_FEATURES",
    "cargo",
    "rustc"
  ),
  value = c(
    unname(Sys.info()[["sysname"]]),
    unname(Sys.info()[["release"]]),
    unname(Sys.info()[["machine"]]),
    R.version.string,
    R.version$platform,
    Sys.getenv("NOT_CRAN", unset = ""),
    Sys.getenv("RUSTFLAGS", unset = ""),
    Sys.getenv("SASSY_RUST_FEATURES", unset = ""),
    paste(safe_system("cargo", "--version"), collapse = "\n"),
    paste(safe_system("rustc", "--version"), collapse = "\n")
  ),
  stringsAsFactors = FALSE
)
machine

rust_verbose <- safe_system("rustc", "-vV")
if (length(rust_verbose) > 0L) {
  cat(paste(rust_verbose, collapse = "\n"), "\n")
}

if (identical(Sys.info()[["sysname"]], "Linux")) {
  lscpu <- safe_system("lscpu")
  if (length(lscpu) > 0L) {
    cat(paste(lscpu, collapse = "\n"), "\n")
  }
  cpuinfo <- "/proc/cpuinfo"
  if (file.exists(cpuinfo)) {
    flags <- grep("^(flags|Features)[[:space:]]*:", readLines(cpuinfo, warn = FALSE), value = TRUE)[1]
    if (!is.na(flags)) {
      simd_flags <- unlist(strsplit(sub("^[^:]+:[[:space:]]*", "", flags), "[[:space:]]+"), use.names = FALSE)
      simd_flags <- grep("^(sse|avx|fma|bmi|neon|asimd)", simd_flags, value = TRUE)
      cat("SIMD-related CPU flags:\n", paste(simd_flags, collapse = " "), "\n", sep = "")
    }
  }
} else if (identical(Sys.info()[["sysname"]], "Darwin")) {
  sysctl <- safe_system("sysctl", c("-n", "machdep.cpu.brand_string"))
  if (length(sysctl) > 0L) {
    cat("CPU: ", paste(sysctl, collapse = " "), "\n", sep = "")
  }
  features <- safe_system("sysctl", c("-a"))
  features <- grep("machdep.cpu.features|machdep.cpu.leaf7_features", features, value = TRUE)
  if (length(features) > 0L) {
    cat(paste(features, collapse = "\n"), "\n")
  }
}
```

## Benchmark configuration

```{r configuration}
pattern <- "AAGGGGA"
unit_text <- "CCCCCCCCCAAGGGGACCCCCAAGGCGACCCCCCCCC"
text <- paste(rep(unit_text, text_multiplier), collapse = "")
raw_pattern <- charToRaw(pattern)
raw_text <- charToRaw(text)
k <- 1L

config <- data.frame(
  batches = batches,
  iterations_per_batch = iterations,
  warmup = warmup,
  text_length = nchar(text, type = "bytes"),
  pattern_length = nchar(pattern, type = "bytes"),
  k = k
)
config
```

## Benchmark

```{r benchmark}
validate <- function(result, label) {
  required <- c("text_start", "text_end", "pattern_start", "pattern_end", "cost", "strand")
  if (!is.data.frame(result) || !all(required %in% names(result))) {
    stop(label, " did not return the expected matches data frame", call. = FALSE)
  }
  if (nrow(result) <= 0L) {
    stop(label, " returned no matches on benchmark input", call. = FALSE)
  }
  invisible(result)
}

bench_one <- function(label, fun) {
  for (i in seq_len(warmup)) {
    validate(fun(), label)
  }

  gc()
  seconds_per_call <- numeric(batches)
  match_count <- NA_integer_

  for (batch in seq_len(batches)) {
    last <- NULL
    elapsed <- system.time({
      for (i in seq_len(iterations)) {
        last <- fun()
      }
    })[["elapsed"]]
    validate(last, label)
    match_count <- nrow(last)
    seconds_per_call[[batch]] <- elapsed / iterations
  }

  data.frame(
    binding = label,
    batches = batches,
    iterations_per_batch = iterations,
    text_length = nchar(text, type = "bytes"),
    pattern_length = nchar(pattern, type = "bytes"),
    k = k,
    matches = match_count,
    min_seconds = min(seconds_per_call),
    median_seconds = unname(stats::median(seconds_per_call)),
    mean_seconds = mean(seconds_per_call),
    p95_seconds = unname(stats::quantile(seconds_per_call, probs = 0.95, names = FALSE)),
    max_seconds = max(seconds_per_call),
    calls_per_second = 1 / unname(stats::median(seconds_per_call)),
    stringsAsFactors = FALSE
  )
}

rsassy_searcher_chr <- Rsassy::sassy_searcher("dna", rc = FALSE)
rsassy_searcher_raw <- Rsassy::sassy_searcher("dna", rc = FALSE)

benchmarks <- list(
  "Rsassy reusable character" = function() {
    Rsassy::sassy_searcher_search(rsassy_searcher_chr, pattern, text, k)
  },
  "Rsassy reusable raw" = function() {
    Rsassy::sassy_searcher_search(rsassy_searcher_raw, raw_pattern, raw_text, k)
  },
  "Rsassy one-shot character" = function() {
    Rsassy::sassy_search(pattern, text, k, alphabet = "dna", rc = FALSE)
  },
  "Rsassy one-shot raw" = function() {
    Rsassy::sassy_search(raw_pattern, raw_text, k, alphabet = "dna", rc = FALSE)
  },
  "sassyRS one-shot character" = function() {
    sassyRS::sassy_search(pattern, text, k, alphabet = "dna", rc = FALSE)
  },
  "sassyRS one-shot raw" = function() {
    sassyRS::sassy_search(raw_pattern, raw_text, k, alphabet = "dna", rc = FALSE)
  }
)

message(
  "Benchmarking ", length(benchmarks), " cases; ", batches, " batches x ",
  iterations, " iterations; text length ", nchar(text, type = "bytes"), " bytes."
)
results <- do.call(rbind, Map(bench_one, names(benchmarks), benchmarks))
row.names(results) <- NULL
results <- results[order(results$median_seconds), ]
results
```

## Summary

```{r summary}
summary_table <- results[, c("binding", "median_seconds", "calls_per_second", "matches")]
summary_table

dir.create(dirname(output), recursive = TRUE, showWarnings = FALSE)
utils::write.csv(results, output, row.names = FALSE)
message("Wrote benchmark results to ", output)

summary_path <- Sys.getenv("GITHUB_STEP_SUMMARY", unset = "")
if (!identical(summary_path, "")) {
  lines <- c(
    "## Rsassy vs sassyRS benchmark",
    "",
    sprintf("Machine: `%s %s (%s)`.", Sys.info()[["sysname"]], Sys.info()[["release"]], Sys.info()[["machine"]]),
    sprintf("Rust: `%s`; cargo: `%s`.", machine$value[machine$field == "rustc"], machine$value[machine$field == "cargo"]),
    sprintf("Build flags: `RUSTFLAGS=%s`; `SASSY_RUST_FEATURES=%s`.", Sys.getenv("RUSTFLAGS", unset = ""), Sys.getenv("SASSY_RUST_FEATURES", unset = "")),
    sprintf("Text length: `%s` bytes; pattern length: `%s`; k: `%s`.", nchar(text, type = "bytes"), nchar(pattern, type = "bytes"), k),
    sprintf("Each row reports the median over `%s` batches of `%s` calls.", batches, iterations),
    "",
    "| Binding | Median seconds/call | Calls/second | Matches |",
    "|---|---:|---:|---:|",
    sprintf(
      "| %s | %.6f | %.2f | %d |",
      results$binding,
      results$median_seconds,
      results$calls_per_second,
      results$matches
    ),
    "",
    sprintf("CSV artifact: `%s`", output)
  )
  writeLines(lines, summary_path, useBytes = TRUE)
}
```
