---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file. -->

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# Rsassy

`Rsassy` provides R bindings to the Rust [`sassy`](https://github.com/RagnarGrootKoerkamp/sassy) approximate string matcher through R's native C API. This package is the lower-level R binding: the C boundary is where raw-vector access and future ALTREP-aware paths can be handled explicitly.

## Installation

From the repository root:

```sh
R CMD INSTALL r/Rsassy
```

For local benchmarking on the current machine, compile with native CPU features:

```sh
RUSTFLAGS="-C target-cpu=native" \
SASSY_RUST_FEATURES=native-simd \
R CMD INSTALL r/Rsassy
```

System requirements are Cargo, `rustc >= 1.91`, and `xz`.

## Usage

```{r}
library(Rsassy)

sassy_search(
  pattern = "ACGT",
  text = "TTACGTAA",
  k = 0,
  alphabet = "dna",
  rc = FALSE
)
```

For repeated searches, create and reuse a searcher:

```{r}
searcher <- sassy_searcher("dna", rc = FALSE)
sassy_searcher_search(searcher, "AAGGGGA", "CCCCCCCCCAAGGGGACCCCCAAGGCGACCCCCCCCC", k = 1)
```

Raw vectors are accepted without converting through R strings:

```{r}
sassy_search(
  charToRaw("ACGT"),
  charToRaw("TTACGTAA"),
  k = 0,
  alphabet = "dna",
  rc = FALSE
)
```

The returned data frame uses 0-based, half-open coordinates: `text_start`, `text_end`, `pattern_start`, and `pattern_end`.

## Testing

`Rsassy` uses `tinytest`:

```sh
Rscript -e 'tinytest::test_package("Rsassy")'
```

## Benchmark

The binding benchmark is in [`../benchmarks/benchmark-r-bindings.Rmd`](../benchmarks/benchmark-r-bindings.Rmd), with a rendered GitHub Markdown report at [`../benchmarks/benchmark-r-bindings.md`](../benchmarks/benchmark-r-bindings.md).
